|

### 2023-01-27 paper.Rmd
Fri Jan 27 15:00:07 CET 2023: update-remote.sh

```{r daegp-flowchart, echo=FALSE, fig.cap="Regular DAE-GP Flowchart", out.height="90%", out.width="50%"}
knitr::include_graphics("./img/flowcharts/dae-gp.png")
```

@@

```{r pt-daegp-flowchart, echo=FALSE, fig.cap="Pre-Trained DAE-GP Flowchart", out.height="90%", out.width="50%"}
knitr::include_graphics("./img/flowcharts/pt-dae-gp.png")
```

@@

```{r airfoil-fullRun-2hl-maxIndSize-params, echo=FALSE, message=FALSE, warning=FALSE}
fig.pos="H"
knitr::kable(
  read.csv("./tables/airfoil_fullRun_2hl_maxIndSize_params"),
  caption = "DAE-GP - Hyperparameter"
)%>% row_spec(0,bold=TRUE )
```

@@

```{r airfoil-function-set, echo=FALSE, message=FALSE, warning=FALSE, paged.print=TRUE}
fig.pos="H"
knitr::kable(
  read.csv("./tables/airfoil_function_set"),
  caption="Airfoil - Function Set"
)%>% row_spec(0,bold=TRUE ) %>% kableExtra::kable_styling(latex_options = "hold_position")
```

<!--- energy heating dataset -->

@@

To study generalization behavior this series of experiments is conducted using DAE-GP with only a single generation until the search terminates. I disregard the fitness of solutions found and focus soley on the reconstruction error that is produced during the training of each the DAE-LSTM. The reconstruction error is measured for two separate populations, a training population $P_{train}$ that is used to train our DAE-LSTM as well as a hold-out validation population $P_{test}$. For the pre-trained DAE-GP two additional, separate populations $\hat{P}_{train}$ and  $\hat{P}_{test}$ are used exclusively for pre-training.

@@

<!--- analyze results 1 hidden layer and variable number of hidden neurons -->

![First Generation Median Training Error for variable number of hidden Neurons - Airfoil](./img/airfoil_firstGen/airfoil_firstGen_mean_training_error_by_neurons_6plots.png)

@@

![First Generation Median Training Error for variable number of hidden Layers - Airfoil](./img/airfoil_firstGen/airfoil_firstGen_mean_training_error_by_layers_6plot.png)

@@

After closely examining the effect of pre-training on DAE-GPs generalization behaviour another series of experiments is conducted to study how pre-training influences the overall search behaviour of DAE-GP. Again the Airfoil problem was selected as a first benchmark problem to test pre-trained DAE-GP against traditional DAE-GP [@dae-gp_2022_symreg], the parameters are described in table

<!-- # TODO: Reformulate -->

@@

![Best Fitness over 30 Generations - Airfoil](./img/airfoil_2hl_maxIndSize_fullRun_30gens_withGP/mean_median_fitness_byGens.png)

@@

To gather more confidence in the previous results, the pre-training strategy for DAE-GP was tested for various real world symbolic regression problems taken from @machine_learning_repo. Are brief summary of the benchmark problems is given in table X

```{r full-run-realWorldSymReg-problems, echo=FALSE, message=FALSE, warning=FALSE, paged.print=TRUE}
fig.pos="H"
knitr::kable(
  read.csv("./tables/full_run_realWorldSymReg_problems.csv"),
  caption="Overview - Real World Symbolic Regression Benchmark Problems"
)%>% row_spec(0,bold=TRUE ) %>% kableExtra::kable_styling(latex_options = "hold_position")
```
Table X summarizes the median best fitness achieved after running each algorithm for 10 runs using a population size of 500 individuals for 30 generations.^[P-Values are marked with asteriks (1,2,3) for ($\alpha$)-levels of $0.1$, $0.05$ and $0.01$]. The median best fitness over all 30 generations are visualized in figure X.

>>>>

```{r daegp-flowchart, echo=FALSE, fig.cap="Regular DAE-GP Flowchart", out.width="60%"}
knitr::include_graphics("./img/flowcharts/dae-gp.png")
```

@@

```{r pt-daegp-flowchart, echo=FALSE, fig.cap="Pre-Trained DAE-GP Flowchart", out.width="65%"}
knitr::include_graphics("./img/flowcharts/pt-dae-gp.png")
```

@@

```{r airfoil-fullRun-2hl-maxIndSize-params, echo=FALSE, message=FALSE, warning=FALSE}
fig.pos="H"
knitr::kable(
  read.csv("./tables/airfoil_fullRun_2hl_maxIndSize_params", sep = "?"),
  caption = "DAE-GP - Hyperparameter"
)%>% row_spec(0,bold=TRUE )
```

@@

Besides studying the performance for the airfoil dataset, additional real world symbolic regressions experiments used the datasets[@machine_learning_repo] described in table \@ref(tab:full-run-realWorldSymReg-problems).

```{r full-run-realWorldSymReg-problems, echo=FALSE, message=FALSE, warning=FALSE}
fig.pos="H"
knitr::kable(
  read.csv("./tables/full_run_realWorldSymReg_problems.csv"),
  caption="Real World Symbolic Regression Benchmark Problems"
)%>% row_spec(0,bold=TRUE ) %>% kableExtra::kable_styling(latex_options = "hold_position")
```

@@

To study generalization behavior this series of experiments is conducted using DAE-GP with only a single generation until the search terminates. I disregard the fitness of solutions found and focus soley on the reconstruction error that is produced during the training of each DAE-LSTM. The reconstruction error is measured for two separate populations, a training population $P_{train}$ that is used to train our DAE-LSTM as well as a hold-out validation population $P_{test}$. For the pre-trained DAE-GP two additional, separate populations $\hat{P}_{train}$ and  $\hat{P}_{test}$ are used exclusively for pre-training.

@@

<!--- analyzing the results for variable number of neurons -->

```{r first-gen-airfoil-byNeurons, echo=FALSE, fig.cap="First Generation Median Training Error for variable number of hidden Neurons", out.width="90%"}
knitr::include_graphics("./img/airfoil_firstGen/airfoil_firstGen_median_training_error_by_neurons_3plots.png")
```

@@

```{r first-gen-airfoil-byLayers, echo=FALSE, fig.cap="First Generation Median Training Error for variable number of hidden Layers", out.width="90%"}
knitr::include_graphics("./img/airfoil_firstGen/airfoil_firstGen_median_training_error_by_layers_3plots.png")
```

@@

After closely examining the effect of pre-training on DAE-GPs generalization behaviour another series of experiments is conducted to study how pre-training influences the overall search behaviour of DAE-GP. The airfoil problem was selected as a first benchmark problem to test pre-trained DAE-GP against traditional DAE-GP.

@@

Figure \@ref(fig:airfoil_median_fitness_30gens) shows both the mean and median of the best fitness for each generation on both the test and training set of the airfoil dataset. The results are aggregated for $10$ individuals runs per algorithm and also include regular GP as a benchmark. For all three algorithms an improvement in the best found fitness can be observed during the evolutionary search but the results show that DAE-GP based algorithms, on average, find solutions with much higher fitness as traditional GP.

The first interesting observation from this experiment is, that the pre-trained DAE-GP shows better performance than regular DAE-GP for both the training and the testing set in overall fitness. Regarding the presence of overfitting, both DAE-GP variants show only a very small gap between the testing and training fitness which indicates a good generalization behaviour.

```{r airfoil_median_fitness_30gens, echo=FALSE, fig.cap="Best Fitness over 30 Generations - Airfoil", out.width="90%"}
knitr::include_graphics("./img/airfoil_2hl_maxIndSize_fullRun_30gens_withGP/mean_median_fitness_byGens.png")
```

@@

To gather more confidence in the previous results, the pre-training strategy for DAE-GP was tested for three additional real world symbolic regression problems taken from @machine_learning_repo.^[see table \@ref(tab:full-run-realWorldSymReg-problems)]

Table X summarizes the median best fitness achieved after running each algorithm for 10 runs using a population size of 500 individuals for 30 generations.^[P-Values are marked with asteriks (1,2,3) for ($\alpha$)-levels of $0.1$, $0.05$ and $0.01$]. The median best fitness over all 30 generations are visualized in figure X.

### 2023-01-27 paper.tex
Fri Jan 27 15:00:07 CET 2023: update-remote.sh

{\centering \includegraphics[width=0.5\linewidth,height=0.9\textheight]{./img/flowcharts/dae-gp}

@@

{\centering \includegraphics[width=0.5\linewidth,height=0.9\textheight]{./img/flowcharts/pt-dae-gp}

@@

\caption{\label{tab:airfoil-fullRun-2hl-maxIndSize-params}DAE-GP - Hyperparameter}
\centering
\begin{tabular}[t]{l|l}
\hline
\textbf{Hyperparameter} & \textbf{Value}\\
\hline
popSize & 500\\
\hline
generations & 30\\
\hline
fitness & RMSE\\
\hline
TrainingMode & Convergence\\
\hline
SamplingSteps & 2\\
\hline
hiddenLayers & 2\\
\hline
Selection & Binary Tournament Selection\\
\hline
corruptionTechnique & Levenshtein Edit\\
\hline
edit\_prob(training/sampling) & 0.05/0.95\\
\hline
functionSet & \{+ ; - ; * ; aq\}\\
\hline
Pre-Training PopulationSize & 10000\\
\hline
Pre-Training Train/Test Split & 50\%\\
\hline
Pre-Training TrainingMode & Early Stopping\\
\hline
\end{tabular}
\end{table}

@@

@@

\caption{\label{tab:airfoil-function-set}Airfoil - Function Set}
\centering
\begin{tabular}[t]{l|r}
\hline
\textbf{operator} & \textbf{arity}\\
\hline
addition & 2\\
\hline
subtraction & 2\\
\hline
multiplication & 2\\
\hline
analytic\_quotient & 2\\
\hline
\end{tabular}
\end{table}

@@

To study generalization behavior this series of experiments is conducted using DAE-GP with only a single generation until the search terminates. I disregard the fitness of solutions found and focus soley on the reconstruction error that is produced during the training of each the DAE-LSTM. The reconstruction error is measured for two separate populations, a training population \(P_{train}\) that is used to train our DAE-LSTM as well as a hold-out validation population \(P_{test}\). For the pre-trained DAE-GP two additional, separate populations \(\hat{P}_{train}\) and \(\hat{P}_{test}\) are used exclusively for pre-training.

@@

\begin{figure}
\centering
\includegraphics{./img/airfoil_firstGen/airfoil_firstGen_mean_training_error_by_neurons_6plots.png}
\caption{First Generation Median Training Error for variable number of hidden Neurons - Airfoil}
\end{figure}

\begin{figure}
\centering
\includegraphics{./img/airfoil_firstGen/airfoil_firstGen_mean_training_error_by_layers_6plot.png}
\caption{First Generation Median Training Error for variable number of hidden Layers - Airfoil}
\end{figure}

@@

To gather more confidence in the previous results, the pre-training strategy for DAE-GP was tested for various real world symbolic regression problems taken from {[}\protect\hyperlink{ref-machine_learning_repo}{2}{]}. Are brief summary of the benchmark problems is given in table X

\caption{\label{tab:full-run-realWorldSymReg-problems}Overview - Real World Symbolic Regression Benchmark Problems}
\centering
\begin{tabular}[t]{l|r|r}
\hline
\textbf{Problem} & \textbf{Observations} & \textbf{Features}\\
\hline
Airfoil & 1503 & 5\\
\hline
Boston\_Housing & 506 & 13\\
\hline
Energy\_Cooling & 768 & 8\\
\hline
Concrete & 1030 & 8\\
\hline
\end{tabular}
\end{table}

>>>>

{\centering \includegraphics[width=0.6\linewidth]{./img/flowcharts/dae-gp}

@@

{\centering \includegraphics[width=0.65\linewidth]{./img/flowcharts/pt-dae-gp}

@@

\caption{\label{tab:airfoil-fullRun-2hl-maxIndSize-params}DAE-GP - Hyperparameter}
\centering
\begin{tabular}[t]{l|l}
\hline
\textbf{Hyperparameter} & \textbf{Value}\\
\hline
Population\_Size & 500\\
\hline
Generations & 30\\
\hline
Fitness\_Metric & RMSE\\
\hline
Training\_Mode & Convergence\\
\hline
Sampling\_Steps & 2\\
\hline
Selection\_Operator & Binary Tournament Selection\\
\hline
Corruption\_Operator & Levenshtein Edit\\
\hline
Denoising\_Edit\_Probability(training/sampling) & 0.05/0.95\\
\hline
Function\_Set & \{+ , - , * , aq\}\\
\hline
Ephemeral\_Constants & [-5,..,5]\\
\hline
Pre-Training\_Population\_Size & 10000\\
\hline
Pre-Training\_Train/Test\_Split & 50\%\\
\hline
Pre-Training\_Training\_Mode & Early Stopping\\
\hline
\end{tabular}
\end{table}

@@

Besides studying the performance for the airfoil dataset, additional real world symbolic regressions experiments used the datasets{[}\protect\hyperlink{ref-machine_learning_repo}{2}{]} described in table \ref{tab:full-run-realWorldSymReg-problems}.

@@

\caption{\label{tab:full-run-realWorldSymReg-problems}Real World Symbolic Regression Benchmark Problems}
\centering
\begin{tabular}[t]{l|r|r}
\hline
\textbf{Problem} & \textbf{Observations} & \textbf{Features}\\
\hline
Airfoil & 1503 & 5\\
\hline
Boston\_Housing & 506 & 13\\
\hline
Energy\_Cooling & 768 & 8\\
\hline
Concrete & 1030 & 8\\
\hline
\end{tabular}
\end{table}

@@

To study generalization behavior this series of experiments is conducted using DAE-GP with only a single generation until the search terminates. I disregard the fitness of solutions found and focus soley on the reconstruction error that is produced during the training of each DAE-LSTM. The reconstruction error is measured for two separate populations, a training population \(P_{train}\) that is used to train our DAE-LSTM as well as a hold-out validation population \(P_{test}\). For the pre-trained DAE-GP two additional, separate populations \(\hat{P}_{train}\) and \(\hat{P}_{test}\) are used exclusively for pre-training.

@@

{\centering \includegraphics[width=0.9\linewidth]{./img/airfoil_firstGen/airfoil_firstGen_median_training_error_by_neurons_3plots}

\caption{First Generation Median Training Error for variable number of hidden Neurons}\label{fig:first-gen-airfoil-byNeurons}
\end{figure}

{\centering \includegraphics[width=0.9\linewidth]{./img/airfoil_firstGen/airfoil_firstGen_median_training_error_by_layers_3plots}

\caption{First Generation Median Training Error for variable number of hidden Layers}\label{fig:first-gen-airfoil-byLayers}
\end{figure}

@@

To gather more confidence in the previous results, the pre-training strategy for DAE-GP was tested for three additional real world symbolic regression problems taken from {[}\protect\hyperlink{ref-machine_learning_repo}{2}{]}.\footnote{see table \ref{tab:full-run-realWorldSymReg-problems}}
